Contrasting influences of topography and canopy height on the functional assembly of canopy and understory tree guilds in an Indian rainforest

### 2020-11-09 Manuscript.Rmd
JVAS formatting

---
title: Topography and canopy height differently shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - orcid.org/0000-0002-9631-114X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - chandra.s.jha@gmail.com
    - National Remote Sensing Centre, Indian Space Research Organisation, Hyderabad 500 037, India 
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - orcid.org/0000-0001-8776-4705
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire d'Écologie Alpine, Univ. Grenoble Alpes, FR-38000 Grenoble, France

@@

# Changes

## 13 octobre 2020

* I changed the title.
* I removed the bidirectional stepwise selection of predictors as it was unstable.
* I used bayesian inference for the linear regression (minor).
* I replaced table 2 by a figure of posteriors distribution and significance (seemed clearer).
* I removed the ideas of LES and WES between strata. That's unclear and difficult to defend. WD model in understorey showed a $R^2=0.3$, which is huge and removed the idea that the WES mainly affects canopy. Similarly, SLA and LDMC models in canopy had low $R^2$ but showed significant effects of wetness, curvature and slope.
* I kept Maxime’s hypothesis for canopy height effect in understory. I don’t have a better hypothesis. I am therefore to keep it calling for further testing. **Maybe we should remove canopy height to simplify the message?**
* I am not a fan of the conclusion. I kept it for the moment but I didn’t feel it brought much more than the take home message; So I’m hesitating to simply remove it.

@@

1. Topographic heterogeneity determines changing soil conditions, while the height of the canopy modulates the microclimate in the forest. But the environmental conditions encountered by trees also change across the vertical strata of tropical forests. However, the joint influence of topography and canopy height with vertical strata on tree functional composition remain little known. We hypothesized different leaf and wood functional strategies between canopy and understory guilds of trees, and different responses of the guilds to topographic heterogeneity and canopy height, depending on resource availability, hydraulic and biomechanical constraints.
1. We measured leaf and wood traits over 89 species representing 99% of all individuals in a 10-ha rainforest plot located in the Western Ghats, South India. The plot was subdivided into tree communities, each composed of guilds of canopy and understory trees. We combined tree census data, LiDAR-derived topographic and canopy metrics, and functional traits to assess the variation of functional composition in understory and canopy guilds, and to address the effects of abiotic and biotic environments on these compositions.
1. The variation of functional traits across sampled tree species followed locally decoupled wood and leaf economics spectra. Although functional space did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
1. *Synthesis.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities. Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

@@

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. 
Environmental variation influences the composition of tree community at regional [@Rejou-Mechain2008; @Tuomisto2003] 
and local scale [@Ferry2007; @gunatilleke_specieshabitat_2006]. 
Topography is known to filter different ecological strategies [@kraft_functional_2008; @Schmitt2020]
due to its impact on water and nutrient availability [@Blanchard2019; @ferry2010higher].
In addition, canopy height modulates forest microclimates due to topography [@Jucker2018].
But, tree species  occupy different vertical strata at adult stage, and can thus be classified into different vertical guilds (understory and canopy),
experimenting contrasted environmental conditions. 
Indeed, light availability decreases downward across vertical forest strata [@chazdon1984],
while hydraulic and mechanical constraints increase upward [@Koch2004], 
so that different survival and growth strategies are filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a]. 
While functional strategies of tree species have been documented independently across topography, canopy height and vertical guilds,
little is known on their interactions,
particularly the effect of topography and canopy height on species assemblages between vertical guilds has not been addressed [@wright2002plant].

Functional traits express fundamental trade-offs determining species occurrences and co-occurrences along environmental gradients [@Wright2002], 
contributing thus to test for community assembly processes. 
Functional traits have been defined as phenotypic traits that impact fitness through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007].
Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018
along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009].
The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016].
Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients. 
The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009].
Previous studies showed decoupled trait variations along the leaf and wood economics spectra for tropical trees at large scale in the Neotropics [@Baraloto2010; @fortunel_leaf_2012]. 
Nevertheless, @reich_world-wide_2014
proposed a unique plant economics spectrum simply opposing fast and slow growing species as observed in subarctic flora [@freschet_evidence_2010].
The concept of plant economics spectrum may be highly relevant for understanding the variation of strategy between vertical guilds as understory species are generally characterized by a low potential growth rate associated to a conservative strategy of resource use, while canopy species tend to have a higher potential growth rate associated to a faster resource acquisition [@Herault2011].

In the context of seasonal rainforests, we hypothesized different trait-environment relationships across vertical guilds, 
with varying strategies along the fundamental leaf and wood economics spectra. 
More specifically, because understory trees are strongly light-limited and are less subject to hydraulic and mechanical constraints than canopy trees, 
we expect that  understory and canopy species tend to maximize their difference along the leaf and wood economic spectra, respectively. 
More generally, we expect that understory species and assemblages have more conservative strategies than canopy ones and that the lower water and nutrient availability found on hilltops [@ferry2010higher] 
entail stronger conservative resource-use strategy [@Cunningham2007; @kraft_functional_2008; @Mendez-Toribio2017; @Schmitt2020]. 
Previous works have shown different response of vertical strata to environmental variation at ground level, with increased edaphic constraint on understory compared to canopy  for Columbian drained uplands [@duque2002different].

Here, we assessed the variation in functional composition of understory and canopy guilds within tropical tree communities, 
and examined the effects of topography and canopy height on these variations. 
We measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot located in the Western Ghats, South India, a region considered as a global biodiversity hotspot [@myers_biodiversity_2000].
The plot encompasses a broad variation of topography [@Antin2013] and vertical structure [@Pascal1988].
Because the region experiences a long dry season period and is under the influence of intense monsoon during rainy seasons with strong wind disturbances, 
we expected hydraulic and mechanical constraints to play an important role in tree community assembly.

We used tree census data, LiDAR-derived topographic and canopy metrics and measured functional traits, 
in multivariate approaches and linear regressions to address the following questions: 
(i) Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale? 
If so, how are canopy and understory guilds distributed along leaf and wood economics spectra? 
(ii) Does functional composition differ between assemblages of canopy and understory guilds? 
and (iii) Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?

@@

The study was carried out in a nearly 10-ha (300 m*330 m) permanent plot established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996]. The plot is named after the nearest village, Uppangala, and is located in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India). Annual precipitation averages 5,100 mm per year but is highly variable across seasons with heavy rainfall during the monsoon from June to October (90% of the rainfall) and a severe dry season (monthly rainfall of 45 mm) lasting 3.8 months on average [@pelissier_twenty_2011]. Mean annual temperature in the station of Sampaji (15 km from Uppangala) is about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [@Pascal1996]. The geologic substrate can be considered homogeneous within the Uppangala plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig. \@ref(fig:Uppangala)). Topography is thus expected to be a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].

```{r Uppangala, fig.cap="Digital canopy model of the 10-ha forest permanent plot in Uppangala, India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
DCM <- raster("./data/Canopy.tif")
crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
DEM <- raster("./data/DEM.tif")
crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
limits <- shapefile("./data/Plots/Shapeall10ha.shp")
limits$melt <- 1
limits <- maptools::unionSpatialPolygons(limits, limits$melt)
india <- shapefile("./data/India_SHP/INDIA.shp")
DCM <- crop(DCM, buffer(limits, 100))
DEM <- crop(DEM, buffer(limits, 100))
DEM <- as(DEM, "SpatialPixelsDataFrame")
DEM <- as.data.frame(DEM)
colnames(DEM) <- c("Elevation", "x", "y")
DCM <- as(DCM, "SpatialPixelsDataFrame")
DCM <- as.data.frame(DCM)
colnames(DCM) <- c("Height", "x", "y")
ggplot() +  
geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
                        stat = "contour", breaks = seq(300, 500, by = 25)) +
  geom_polygon(data = fortify(limits),
               aes(long, lat, group=group),
               colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  cowplot::theme_map() +
  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
                 transform = F, model = "WGS84", location = "bottomleft") +
    annotation_custom(grob = ggplotGrob(ggplot() +  
                                        geom_polygon(data = fortify(india),
                                                     aes(long, lat, group=group),
                                                     colour = "black", alpha = 0) +
                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
                                                   aes(long, lat), col = "red", size = 3) +
                                        coord_equal() +
                                        cowplot::theme_map() +
                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
                                              plot.background	= element_rect(fill = "white"))), 
                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
```

## Taxonomic composition and trait sampling

@@

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species corresponding to the vertical stratum reached by most adult trees.
The strata included four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. 
Because our study site is characterized by a steep topography (Fig. \@ref(fig:Uppangala)), 
sub-canopy species can often reach the canopy in the Uppangala plot. 
We thereby defined two broader categories, namely, canopy guild (C) including emergent, 
canopy and sub-canopy species and understory guild (U) composed of understory trees.

We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species). We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance. For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following the recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; [@Dawkins1958]). Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density.

## Trait measurements

We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002]: leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$). Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weight was measured with an analytical balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness with a micrometer and leaf area by  analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012]. Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight. The bark and pit of branch samples were removed in order to keep only branch wood without soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. Wood samples were then dried for 48 hours at 80°C and weighed following the protocol of @Perez-Harguindeguy2013.

@@

if(knitr:::is_latex_output()){
  t %>% 
    ungroup() %>% 
    mutate(Trait = c('Leaf area', 'Leaf dry\nmatter content', 'Leaf\nthickness', 'Specific\nleaf area', 'Wood\ndensity'),
    Strategy = c('Leaf defense\nand investment',
                 'Leaf resource\ncapture vs. investment',
                 'Leaf defense\nand investment',
                 'Leaf resource\nacquisition vs. defense',
                 'Stem transport,\nstructure and defense')) %>% 
    mutate_all(linebreak) %>% 
    kable(caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).',
          digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
          col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "\\makecell[l]{Intra-\\\\individuals}", 
                        "\\makecell[l]{Intra-\\\\species}", 
                        "\\makecell[l]{Inter-\\\\species}")) %>%
    kable_styling(full_width = T)
} else {  
  t %>% 
  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
                  col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "Intra-individuals", "Intra-species", "Inter-species"),
        caption = 'Functional traits measured. Traits are presented with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).') %>% 
  kable_styling(full_width = T)
}
```

@@

To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific). We then performed a principal component analysis (PCA) of average trait values per species to characterize the main dimensions of ecological strategies globally as well as per vertical stratum.

We subsampled tree communities in $30~m*30~m$ subplots (n=110), in order to characterize the functional composition of vertical guilds within each subplot. Each subplot contained 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds. We thus calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) for each trait in each subplot:

>>>>

---
title: Contrasting influences of topography and canopy height on the functional assembly of canopy and understory tree guilds in an Indian rainforest
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/journal-of-vegetation-science.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

@@

1. Sylvain Schmitt
    - Corresponding author
    - sylvain.m.schmitt@gmail.com
    - orcid.org/0000-0001-7759-7106
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - UMR BIOGECO, INRA - Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
1. Valérie Raevel
    - vraevel@gmail.com
    - orcid.org/0000-0002-9631-114X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Maxime Réjou-Méchain
    - maxime.rejou@ird.fr
    - orcid.org/0000-0003-2824-267X
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - AMAP, IRD, CNRS, CIRAD, INRA, Univ Montpellier, Montpellier, France
1. Narayanan Ayyappan
    - ayyappan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Natesan Balachandran
    - balachandran.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. N. Barathan
    - barathan.n@ifpindia.org
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
1. Chandra Shekhar Jha
    - chandra.s.jha@gmail.com
    - National Remote Sensing Centre, Indian Space Research Organisation, Hyderabad 500 037, India 
1. François Munoz
    - fmunoz@univ-grenoble-alpes.fr
    - orcid.org/0000-0001-8776-4705
    - French Institute of Pondicherry, UMIFRE 21/USR 3330 CNRS-MAEE, Pondicherry, India
    - Laboratoire Interdisciplinaire de Physique, Univ. Grenoble Alpes, FR-38000 Grenoble, France

**Running title:** Functional assembly of canopy and understory tree

# Funding

No funding to declare. **Comment: ?**

@@

@@

* *Questions.* Topographic heterogeneity determines changing soil conditions, while canopy height modulates the microclimate in forest ecosystems. In addition, tree species form multiple vertical layers undergoing different light, biomechanic, and hydraulic constraints. However, how topography and canopy height influence the functional composition of vertical strata remain little addressed. We hypothesized that leaf and wood functional strategies differ between canopy and understory tree assemblages depending on local environmental heterogeneity.
* *Location.* An undisturbed, old-growth wet evergreen Dipterocarp forest plot located in the Kadamakal Reserve Forest, Karnataka State, India.
* *Methods.* We measured leaf and wood traits of 89 tree species representing 99% of all individuals in a 10-ha rainforest plot in the Western Ghats, South India. We subdivided the plot into subplots representing varying topographic conditions. We assigned tree species to guilds of canopy and understory trees. We combined tree census data, LiDAR-derived topographic and canopy metrics to assess the variation of the weighted means of functional trait values in understory and canopy guilds with canopy height and topography.
* *Results.* The variation of functional traits across sampled tree species followed decoupled wood and leaf economics spectra at the local scale. Although functional trait values did not differ overall between canopy and understory tree species, environmental filtering shaped different functional composition of canopy and understory guilds in tree communities. Furthermore, the functional composition of the canopy and understory guilds responded differently across communities to environmental gradients related to soil resource, light availability and aspect.
* *Conclusions.* These results show that both fine-scale environmental heterogeneity and vertical structure shape the composition of hyperdiverse rainforest tree communities. Guild-specific responses to biotic and abiotic environmental filtering underline that vertical stratification must be acknowledged to better understand and forecast forest dynamics.

@@

Understanding the drivers of tree composition in hyperdiverse tropical forests is a long-standing issue in ecology [@Cody1976]. Environmental variation influences the composition of tree communities at regional [@Rejou-Mechain2008; @Tuomisto2003] and local [@Ferry2007; @gunatilleke_specieshabitat_2006] scales. Topography influences water and nutrient availability [@Blanchard2019; @ferry2010higher], in interaction with canopy height [@Jucker2018], and thereby filters different ecological strategies [@kraft_functional_2008; @Schmitt2020]. In addition,  tree species form different vertical strata at the adult stage allowing to distinguish guilds of understory and canopy trees. These vertical guilds should undergo contrasting environmental conditions: light availability decreases downward across vertical forest strata [@chazdon1984], while hydraulic and mechanical constraints increase upward [@Koch2004]. Therefore, we expect different survival and growth strategies to be filtered across strata [@Raevel2018; @Stark2015; @Uriarte2004a]. While much emphasis has been put on how ecological strategies of tree species vary with topography or canopy height in forests, far fewer works have addressed whether and how these variations depend on vertical guilds [@wright2002plant].

Functional traits have been defined as phenotypic traits that impact fitness through their effect on individual performance, which is defined as the ability to recruit, grow, survive, and reproduce in a given environment [@violle_let_2007]. Functional traits covary among species [@diaz_global_2015] and communities [@Bruelheide2018] along distinct economics spectra of leaf [@wright_worldwide_2004; @osnas_global_2013; but see @lloyd_photosynthetically_2013] and wood [@chave_towards_2009]. The leaf economics spectrum expresses a trade-off between resource acquisition by photosynthesis and investment in leaf defense and durability [@Evans2001a; @hodgson_is_2011; but see @Gibert2016]. Trait values along the leaf economics spectrum are expected to change along soil and light availability gradients, with more acquisitive strategies with decreasing resources availability [@hodgson_is_2011]. The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species The wood economics spectrum opposes slow growing, drought-resistant species to fast growing, drought-intolerant species [@chave_towards_2009]. Whether and how trait variations along the leaf and wood economics spectra are decoupled across tropical tree species remain debated [@Baraloto2010; @fortunel_leaf_2012; but see @reich_world-wide_2014; @freschet_evidence_2010].

Here, we measured leaf and wood traits of 89 species representing 99% of all individuals in a 10-ha forest plot. We categorized tree species as belonging to canopy and understory guilds at the adult stage. First, we analyzed the variation of functional traits across species and between guilds. Second, we combined trait information with tree census data, LiDAR-derived topographic and canopy metrics to characterize the functional composition of tree assemblages and their variation with topography [@Antin2013] and vertical structure [@Pascal1988]. We used multivariate approaches and linear regressions to address the following questions:

* Do functional trait variations among tree species follow independent leaf and wood economics spectra at local scale? If so, how are canopy and understory guilds distributed along leaf and wood economics spectra? 
* Does functional composition differ between assemblages of canopy and understory guilds? 
* Does the functional composition of guild assemblages vary with local environmental heterogeneity, and do the understory and canopy assemblages respond differently?

We hypothesized different trait-environment relationships across vertical guilds, with varying strategies along the leaf and wood economics spectra. First we expected understory species to be strongly light-limited but less subject to hydraulic and mechanical constraints than canopy species, so that understory species should have more conservative strategy [@herault_functional_2011] but less drought tolerance than canopy species . Second, because of the long dry season period and influence of intense monsoon during rainy seasons with strong wind disturbances, we expected topography and canopy height to influence particularly resource use strategies in our plot. Lower water and nutrient availability found on hilltops [@ferry2010higher] should entail more conservative strategy [@Cunningham2007; @kraft_functional_2008;@Schmitt2020], while acquisitive strategies are expected in bottom-lands.

@@

The nearly 10-ha (300 m*330 m) permanent Uppangala plot was established in 1990 in an undisturbed, old-growth wet evergreen Dipterocarp forest [@Pascal1996] in the Kadamakal Reserve Forest (Karnataka State, India). Annual precipitation averages 5,100 mm per year but is highly variable across seasons with most rainfall (90%) occurring during the monsoon from June to October,  while a severe dry season (monthly rainfall of 45 mm) lasts 3.8 months on average [@pelissier_twenty_2011]. Mean annual temperatureis about 27°C, ranging from 25°C at rainfall peak to 29°C at dry season peak [data from a station at 15 km; @Pascal1996]. The geologic substrate is homogeneous within the plot [@Ferry1994], but the topography is highly heterogeneous with regular east-west alternation of thalwegs and ridges separated by steep slopes (Fig. \@ref(fig:Uppangala)). Topography is thus expected to be a major component of environmental variation within the forest plot [@Antin2013; @gimaret-carpentier_sampling_1998].

```{r Uppangala, fig.cap="Digital canopy model built from LiDAR data of the 10-ha forest permanent Uppangala plot , India. Thin and thick black lines represent the elevation contour lines and plot limits, respectively."}
DCM <- raster("./data/Canopy.tif")
crs(DCM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
DEM <- raster("./data/DEM.tif")
crs(DEM) <- '+proj=utm +zone=43 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0'
limits <- shapefile("./data/Plots/Shapeall10ha.shp")
limits$melt <- 1
limits <- maptools::unionSpatialPolygons(limits, limits$melt)
india <- shapefile("./data/India_SHP/INDIA.shp")
DCM <- crop(DCM, buffer(limits, 100))
DEM <- crop(DEM, buffer(limits, 100))
DEM <- as(DEM, "SpatialPixelsDataFrame")
DEM <- as.data.frame(DEM)
colnames(DEM) <- c("Elevation", "x", "y")
DCM <- as(DCM, "SpatialPixelsDataFrame")
DCM <- as.data.frame(DCM)
colnames(DCM) <- c("Height", "x", "y")
ggplot() +  
geom_tile(data = DCM, aes(x = x, y = y, fill = Height),  alpha = 0.6) +
  geom_contour(data = DEM, aes(x = x, y = y, z = Elevation),
               colour = "black", alpha = 0.5, breaks = seq(300, 500, by = 25)) +
  directlabels::geom_dl(data = DEM, aes(x = x, y = y, z = Elevation, label = paste0(..level.., "m")),
                        method = list('last.bumpup', cex = 0.8, hjust = 1, fontface = 2),
                        stat = "contour", breaks = seq(300, 500, by = 25)) +
  geom_polygon(data = fortify(limits),
               aes(long, lat, group=group),
               colour = "black", lwd = 1.1, alpha = 0) +
  coord_equal() +
  cowplot::theme_map() +
  scale_fill_distiller("Canopy\nheight (m)", palette = "Spectral") +
  ggsn::scalebar(dist = 50, dist_unit = "m", x.min = min(DCM$x) + 5, x.max = max(DCM$x), y.min = min(DCM$y) + 20, y.max = max(DCM$y),
                 transform = F, model = "WGS84", location = "bottomleft") +
    annotation_custom(grob = ggplotGrob(ggplot() +  
                                        geom_polygon(data = fortify(india),
                                                     aes(long, lat, group=group),
                                                     colour = "black", alpha = 0) +
                                        geom_point(data = data.frame(long = 75.662778, lat = 12.537500), 
                                                   aes(long, lat), col = "red", size = 3) +
                                        coord_equal() +
                                        cowplot::theme_map() +
                                        theme(panel.border = element_rect(color = "black",size = .5,fill = NA),
                                              plot.background	= element_rect(fill = "white"))), 
                    xmin = min(DCM$x), xmax = min(DCM$x) + 150, 
                    ymin = max(DCM$y) - 150, ymax = max(DCM$y))
  # ggsave(file = "Plot.png", path = './data/', width = 150, height = 200, unit = 'mm', dpi = 300)
```

## Census data and vertical guilds

@@

Using the classification of @Pascal1988, we assigned a potential vertical stratum to each tree species corresponding to the vertical stratum reached by the adult trees. The strata included four categories: emergent above 39 m, canopy between 25 and 39 m, sub-canopy between 15 and 25 m, and understory below 15 m. Because our study site is characterized by a steep topography (Fig. \@ref(fig:Uppangala)), sub-canopy species can often reach the canopy in the Uppangala plot. We thereby defined two broader categories, namely, canopy guild (C) including emergent, canopy and sub-canopy species and understory guild (U) composed of understory trees.

## Functional trait measurements

We collected trait data for 89 over the 111 (80%) species, representing 99% of tree individuals in the plot (54 canopy and 35 understory species). We sampled five individuals per species for the 32 most abundant species (>80% of individuals), and one to five individuals for the remaining species (3 individuals on average), depending on leaf accessibility and species abundance. For each individual, we collected at 4 to 12 m height five mature leaves and one branch sample with a diameter between 1 and 2 cm, following the recommendation of @Perez-Harguindeguy2013. Whenever possible, individuals and leaves were selected in similar light conditions (Dawkins index of 2; @Dawkins1958) **comment: beware double space**. Leaf and branch samples were kept in humidified ziplock bags filled with enriched $CO_2$ air in dark until measurement within the next 24 hours. For 11 species (1% of individuals in Uppangala plot), branch samples were too small (< to 1 cm in diameter) to allow measuring branch wood density.

We measured five functional traits (Table \@ref(tab:traits)) related to the major dimensions of ecological strategies in plants [@diaz_global_2015; @Wright2002] following the protocol of @Perez-Harguindeguy2013: 
leaf thickness (LT, $\mu m$), leaf area (LA, $mm^2$), leaf dry matter content (leaf dry weight divided by the leaf fresh weight, LDMC, $mg.g^{-1}$), specific leaf area (leaf area divided by the leaf dry mass, SLA, $m^2.kg^{-1}$), and branch wood density (wood dry weight divided by its volume, WD, $g.cm^{-3}$).
Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weight was measured with an digital balance at a 0.01 g precision (Wensar, KD-H600), leaf fresh thickness measured with a micrometer and leaf area by analyzing scanned images of fresh leaves with the ImageJ software [@Schneider2012]. Leaves were then kept in herbarium and oven-dried for at least 72 hours at 60°C before measurement of dry weight. We removed the bark and pit of branch samples in order to keep only branch wood without soft tissues. Branch wood samples were immersed in water to evaluate their volume by water displacement. We dried wood samples for 48 hours at 80°C before weighing.

@@

if(knitr:::is_latex_output()){
  t %>% 
    ungroup() %>% 
    mutate(Trait = c('Leaf area', 'Leaf dry\nmatter content', 'Leaf\nthickness', 'Specific\nleaf area', 'Wood\ndensity'),
    Strategy = c('Leaf defense\nand investment',
                 'Leaf resource\ncapture vs. investment',
                 'Leaf defense\nand investment',
                 'Leaf resource\nacquisition vs. defense',
                 'Stem transport,\nstructure and defense')) %>% 
    mutate_all(linebreak) %>% 
    kable(caption = "Functional traits measured, with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).",
          digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
          col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "\\makecell[l]{Intra-\\\\individuals}", 
                        "\\makecell[l]{Intra-\\\\species}", 
                        "\\makecell[l]{Inter-\\\\species}")) %>%
    kable_styling(full_width = T)
} else {  
  t %>% 
  kable(digits = 2, format.args = list(big.mark = " "), format = 'pandoc', escape = F,
                  col.names = c("Trait", "Abbreviation", "Unit", "Strategy", "Range", 
                        "Intra-individuals", "Intra-species", "Inter-species"),
        caption = "Functional traits measured, with their abbreviation, unit, related species strategy according to @baraloto_decoupled_2010, range of values, intra-individual, intra-specific and inter-specific coefficients of variation (CV).") %>% 
  kable_styling(full_width = T)
}
```

@@

To assess the degree of trait variation and covariation, we first calculated the coefficient of variation (ratio of the standard deviation to the mean) of trait values among samples within individuals (intra-individual), among individuals within species (intra-specific) for species with more than one individual, and among species (inter-specific). We then performed a principal component analysis (PCA) of the average trait values per species to characterize the main dimensions of ecological strategies. We compared trait values between species according to the different vertical guilds.

We subsampled tree communities in $30~m*30~m$ subplots (n=110), and characterized the functional composition of vertical guilds within each subplot. Each subplot included 28 individuals and 10 species on average, with less species in understory (6 in average) than in canopy (13 in average) guilds. We calculated the weighted mean trait value per guild (GWM, where G stands for ‘guild’; [@Raevel2018]) **comment: beware double space** for each trait in each subplot:
